Bayesian Networks for Assessing Environmental Sustainability in Australian Vineyards

### 2024-08-29 article.tex
Mardi edits done - a couple lookups left

The imperative for sustainability in agriculture has intensified as the sector strives to balance productivity with ecological preservation~\cite{baianoOverviewSustainabilityWine2021,corboEnvironmentalSustainabilityPrograms2014}. This is especially true for viticulture, the cultivation of grapevines for wine production, as it presents unique challenges and opportunities for sustainability, particularly due to its position as a luxury good amongst other agricultural industries~\cite{longbottomExploringLinksSustainability2018,ferraraLifeCycleAssessment2018}. This study investigates sustainability practices in Australian winegrowing, using the application of Bayesian Networks (BNs) to create a broad model of Australian winegrowing. This model was followed by a more specific case study looking at the Coonawarra winegrowing region. Due to the diverse nature of Sustainable practices and the complicated interactions with environmental, social and economic outcomes~\citep{beaumelle_biodiversity_2023} criterion for assessing sustainability has become multifaceted and led to the creation of many institutions with varying industry-specific frameworks for sustainability metrics (Baiano, 2021; Montalvo-Falcón et al., 2023; Waye, 2008). In this study we focus on the environmental aspect of sustainability through the use of 'environmental impact' as the terminal node within the network, linking relevant factors through structures described by an expert panel.
% TODO: insert reference above
BNs are probabilistic models that capture and analyse the interactions between multiple factors in complex systems~\cite{korbBayesianArtificialIntelligence2011}. BNs also facilitate the integration of diverse types and sources of information, including expert knowledge, the results of prior experiements, outcomes from published literature, and so on. In this study, the BNs integrated expert knowledge and region-specific data within a structured framework to quantify the impacts of various factors on sustainability within Australian winegrowing. This methodology shed light on current sustainability practices, identified critical areas for intervention and facilitated evolution of the potential impact of various management action or climate scenarios.

@@

The literature review was designed to access existing knowledge on sustainability in viticulture with the aim of identifying key factors influencing sustainability and understanding their interactions.

These findings were combined with the information derived from the expert consultations to inform the BN construction and quantification. 
% TODO: Clarify that the review helped inform the strawman model and was only used as a comparison point afterwards.

A systematic search of terms, broadly based on key factors given from the industry panel, was conducted. The general query (adapted to different formats for various journals and bibliographic databases) used for the literature search was:

@@

% How can these key factors be rigorously defined so that they are quantifiable contributions to sustainability?
% Purpose: Develop the factors into well-defined and measurable impacts on sustainability.
% Results: Use the agreed upon definitions to fully realise factors and their impact.
% 
% Please add the following required packages to your document preamble:
% \usepackage{booktabs} TODO: check the BN definitions align properly.
%

@@

    \textbf{Climate Conditions} & A vineyard’s regional climatic conditions. & \textbf{Ideal} & Not Ideal \\

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{Winegrowing-concept-map}
    \caption{Your caption here}
    \label{fig:lexi}
\end{figure}

A key difference between the concepts from the literature search and the BN based on the industry panel was the rarity in how secondary and tertiary drivers of industry such as fuel or electricity use, were connected to primary sustainability concerns such as disease (even though disease was quite prominent in both).\
% TODO: THis is a key point can you explore this a bit more by telling us the practical impact and implications? DOes it mean the prev research has focussed on the wrong areas and what does it mean we need to do? More in a particular area?
% TODO: Note the discussion with Mardi surrounding causality. And discuss further the differences later on about how the models both result in the same significant concerns for sustainability even though they are different models and even though some of the causality is shifted.
%  What does it mean? That some operational decisions are paramount to sustainability in an environmental context.
% That different approaches can have different causal relationships - i.e spraying for prevention and spraying in response to disease.
% That management strategies might change the emphasis of certain affects, or change the ability to control different minor outcomes but some concerns are more important and require direct attention such as disease pressures.
% It is incredibly important that we can acknowledge that some of these relationships can exist as dual casualities or bi-directionally, which brings into question models that cannot take this into account. Or take into account the difference in management strategies and how that does or does not change environmental concerns.

% With these concepts being related to the operations in fields as a result of changing practices or implementing different pest controls. Together these outputs provide a comprehensive overview of the key factors affecting sustainability in the Australian winegrowing industry, offering valuable insights into areas of focus and potential improvement. This analysis underscores the multifaceted nature of sustainability, highlighting both the primary concerns and the interconnected secondary factors that must be managed on the ground.
% TODO: ML: I don't understand how these words were selected. Are they concepts or just frequently used words. Are they al relevant? May need some explanation. - in reference to the Leximancer diagram.

@@

Two distinct BNs were constructed for the Coonawarra region, each reflecting alternative points of view taken by experts. The models diverged in their structural representations with a differing emphasis on specific causal relationships. One BN  (see Figure \ref{fig:marcel} explicitly incorporated connections between soil degradation and tractor passes, as well as between climate conditions and agrochemical scores. The other BN, while structurally similar, did not include these connections. The use of two separately scored BNs with differing structure ensured that the BNs captured the diversity of expert views regarding the primary drivers of environmental outcomes in the region.

@@

    %     Sustainability in Coonawarra Winegrowing: Key Factors and Influences
    % #TODO: what are the regional specifics of coonawarra
    % How do the identified key factors vary across different wine regions in Australia, and what regional characteristics influence these variations?
    %     Purpose: Examine regional differences and tailor sustainability strategies.
    %     Results: Compare the generic model to the Coonawarra model. Acknowledge there is more work to be done here due to the amount of regions across Australia and the world. Prevous work also helps to illustrate these differences.

@@

% TODO: the exclusion of renewable fuels due to none being used in coonawarra
% the difference in vineyard floor management and its relationship to tractors
% The influence of compost, soil depth and clay content. although not specifically significant due to not listing the CPTs - it was highlighted as significant with different influence values depending on the direct field in question and not just which vineyard.
% variation occurs all the way to the vine level!

@@

The primary difference in interactions between the two BNs (see Figure~\ref{fig:robyn} and~\ref{fig:marcel}) lies in vineyard floor management, seasonal conditions and tractor passes. In Figure \ref{fig:marcel}, a higher number of tractor passes were used purposefully to manage the vineyard floor; as opposed to Figure \ref{fig:robyn} where vineyard floor was managed as a consequence of tractor passes, leveraging opportunities when the tractor was used. 
% TODO: the prior sentance is not likely to be understood by a more general audience. also why?
This distinction highlighted how operational decisions can vary based on regional practices and priorities. A point was also raised that scale can cause a further driving factor, where tractor passes may be greatly increased as an expense.
% TODO: This requires further explanation.
However, this issue as well as directly attributing measurable cause to granular specifics such as composting, soil depth and clay content cannot be fully addressed without site or paddock specific knowledge.
% TODO: The other had fewer tractor passes because they performed 2 operations in a single pass. 
% ML: As I read this, I wander whether there was some bias because the group were very attuuned to the data and had a strong preference to efficient management techniques. Im not convinced that this is representative of coonawarra per se.

As a factor, water emerged as the most interconnected variable for both BNs, with edges connecting it to almost everything either directly or indirectly, illustrating its fundamental role in agriculture. Climate conditions also play a significant role, influencing water availability, quality and disease/pest pressures.
% TODO: This is significant...
Within both of the models, the most significant factors impacting sustainability are water quality, pest and disease management, market pressures, agrochemical use, and climate conditions. These factors were identified as having the highest sensitivity to environmental impact, indicating their crucial roles in the sustainability of the region.

@@

The Coonawarra-specific BNs both support the findings from the literature, with a more structured map of how factors interact compared to the Leximancer summary in Figure \ref{fig:lexi}. The BN provides a detailed view of the relationships between factors, highlighting the central role of water and the interconnectedness of pest management, market pressures, and climate conditions. Additionally, the BNs emphasise economic factors, such as market pressures and grape prices, and includes how they interact within a scope of secondary factors like fuel and electricity use.
% TODO: Highlight, in the above paragraph, how the BN differs from other BNs

@@

\section{Discussion}
% weather or climate?
This investigation highlights the key factors of Australian winegrowing that contribute to  sustainability through the lens of BNs created with assistance from industry input. These models have established the intricate, and interconnected nature of various influences on sustainability in Australian winegrowing. A keener understanding of how these factors both interact and affect one another from the perspective of research and industry is illustrated within these models, with a major outcome being the highlighting of water use's critical importance as one of the most interconnected variables.
%TODO: And therefore what to sustainability?
The influence of water use reached almost every variable, with its connection to operational decisions being reflected within its propagation to other factors such as tractor use, disease and pest prevention, as well as management strategies for soil. This finding aligns with water being the lifeblood of agriculture~\cite{chawlaWaterProductivityAgriculture2023}, essential for both the health of the vines and the broader ecosystem, especially within Australia's dry climate~\cite{australianbureauofstatisticsWaterUseAustralian2021}.

@@

The BN models helped to illustrate how economic incentives can be drivers of sustainable practices. The definitive influence and relationships between what are more tertiary considerations of sustainability being directly connected to further considerations such as operational costs. Market pressures, including the demand for high-quality grapes through factors such as grape prices, were found to have an impact on sustainability practices but were limited in this study through only reviewing them within the lens of sustainability. The ability to effectively gauge whether operations could afford to engage further with management programs for factors such as soil, pests, disease etc was in part driven by the economic climate provided to growers as well as what may be much more specific circumstance; of which is a potential further avenue of research that would compliment this work significantly.
% TODO: What do you mean here?

>>>>

The imperative for sustainability in agriculture has intensified as the sector strives to balance productivity with ecological preservation~\cite{baianoOverviewSustainabilityWine2021,corboEnvironmentalSustainabilityPrograms2014}. This is especially true for viticulture, the cultivation of grapevines for wine production, as it presents unique challenges and opportunities for sustainability, particularly due to its position as a luxury good amongst other agricultural industries~\cite{longbottomExploringLinksSustainability2018,ferraraLifeCycleAssessment2018}. This study investigates sustainability practices in Australian winegrowing, using the application of Bayesian Networks (BNs) to create a broad model for Australian. This broad model was followed by a more specific case study looking at the Coonawarra winegrowing region. Due to the diverse nature of Sustainable practices and the complicated interactions with environmental, social and economic outcomes~\citep{beaumelle_biodiversity_2023} criterion for assessing sustainability has become multifaceted and has led to the creation of many institutions with varying industry-specific frameworks for sustainability metrics \citep{baiano_overview_2021,montalvo-falconSustainabilityResearchWine2023, waye_carbon_2008}. In this study we focus on the environmental aspect of sustainability through the use of 'environmental impact' as the terminal node within the network, linking relevant factors through structures described by an expert panel.

BNs are probabilistic models that capture and analyse the interactions between multiple factors in complex systems~\citep{korbBayesianArtificialIntelligence2011}. BNs also facilitate the integration of diverse types and sources of information, including expert knowledge, the results of prior experiements, outcomes from published literature, and so on. In this study, the BNs integrated expert knowledge and region-specific data within a structured framework to quantify the impacts of various factors on sustainability within Australian winegrowing. This methodology shed light on current sustainability practices, identified critical areas for intervention and facilitated evolution of the potential impact of various management action or climate scenarios.

@@

The literature review was designed to assess existing knowledge on sustainability in viticulture with the aim of identifying key factors influencing sustainability and understanding their interactions. These findings were combined with the information derived from the expert consultations to inform the BN construction and quantification. The most prominent way of informing the model was the construction of an initial strawman model to describe different relationships explored in the literature, which helped as a beginning point for the experts to discuss the specific applicability of relationships for different Australian contexts. The literature review was done using systematic search of terms, broadly based on key factors given from the industry panel, was conducted. The general query (adapted to different formats for various journals and bibliographic databases) used for the literature search was:

@@

% How can these key factors be rigorously defined so that they are quantifiable contributions to sustainability?
% Purpose: Develop the factors into well-defined and measurable impacts on sustainability.
% Results: Use the agreed upon definitions to fully realise factors and their impact.
% 
% Please add the following required packages to your document preamble:
% \usepackage{booktabs}
%

@@

    \textbf{Climate Conditions} & A vineyard's regional climatic conditions. & \textbf{Ideal} & Not Ideal \\

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{Winegrowing-concept-map}
    \caption{TODO: add a caption the explains what the connection between concepts in lexi is and how words are selected and culled - what is a concept. Reference the Lexi proceedures followed too}
    \label{fig:lexi}
\end{figure}

A key difference between the concepts from the literature search and the BN based on the industry panel was the rarity in how secondary and tertiary drivers of industry such as fuel or electricity use, were connected to primary sustainability concerns such as disease (even though disease was quite prominent in both). Notably the sensitivity of the Coonawarra models align with the focus of literature and analyses, focusing on  aspects such as disease, water, soil health and climate. The combination of agreeing with these primary concerns whilst highlighting their connection to operational variables such as tractor, and fuel use; is an important step to realising the benefits of combined efforts of prevention through scenario assessments. Considering variables that are related to an operational standpoint as well as environmental Furthers the ability to show the consideration of environmental impact within vineyard operational decisions and how it may lower the costly use of agrochemicals and fuel.

% With these concepts being related to the operations in fields as a result of changing practices or implementing different pest controls. Together these outputs provide a comprehensive overview of the key factors affecting sustainability in the Australian winegrowing industry, offering valuable insights into areas of focus and potential improvement. This analysis underscores the multifaceted nature of sustainability, highlighting both the primary concerns and the interconnected secondary factors that must be managed on the ground.

@@

Two distinct BNs were constructed for the Coonawarra region, each reflecting alternative points of view taken by experts. The models diverged in their structural representations with a differing emphasis on specific causal relationships. One BN (see Figure~\ref{fig:marcel} explicitly incorporated connections between soil degradation and tractor passes, as well as between climate conditions and agrochemical scores. The other BN (see \ref{fig:robyn}), while structurally similar, did not include these connections. The node for renewable fuels was also removed in Figure~\ref{fig:robyn} to more accurately reflect the lack of use of them in the Coonawarra region, whereas Figure~\ref{fig:marcel} included the node to illustrate the potential benefits if it were used. The differences in structure reflect difference in management approaches and the way problems can be approached. This difference in management style was most apparent in the causal relationship between tractor passes and undervine management, where the edge between the nodes is reversed between the two models. The difference highlights how operational decisions can alter the causal relationship of nodes to one another, with an example discussed regarding the use of agrochemical spray as preventative vs as a response to disease. The first decision would have the act of spraying cause a lower probability of disease to occur, and the second being that the presence of disease would be more likely to initiate the use of agrochemicals. The two arguments for the relationship between tractor passes and under vine management reflect this type of causal reversal, where they differed in the use of tractors to actively manage the undervine area or to leverage the tractor when it was already being used for other tasks to then manage the undervine as a consequence. These types of differences in management perspectives can greatly effect the root decision that impacts the environmental even if they do not directly change the most sensitive or contributing factors, as in this case both models maintained similar sensitivities to agrochemical use, water, and climate.

@@

    %     Sustainability in Coonawarra Winegrowing: Key Factors and Influences
    % How do the identified key factors vary across different wine regions in Australia, and what regional characteristics influence these variations?
    %     Purpose: Examine regional differences and tailor sustainability strategies.
    %     Results: Compare the generic model to the Coonawarra model. Acknowledge there is more work to be done here due to the amount of regions across Australia and the world. Prevous work also helps to illustrate these differences.

@@

% the difference in vineyard floor management and its relationship to tractors
% The influence of compost, soil depth and clay content. although not specifically significant due to not listing the CPTs - it was highlighted as significant with different influence values depending on the direct field in question and not just which vineyard.
% variation occurs all the way to the vine level!

@@

The primary difference in interactions between the two BNs (see Figure~\ref{fig:robyn} and~\ref{fig:marcel}) lies in vineyard floor management, seasonal conditions and tractor passes. In Figure \ref{fig:robyn}, a higher number of tractor passes were used purposefully to manage the vineyard floor; as opposed to Figure \ref{fig:marcel} where vineyard floor was managed as a consequence of tractor passes, leveraging opportunities when the tractor was used. A distinction which highlighted how operational decisions and priorities can vary and influence the cause of environmental impacts without changing the involved factors, simply through changing the 'why' of nodes' connections. A point was also raised that scale can cause a further driving factor. The data helped to inform the level of influence scale can have with many vineyards being anywhere from a few to hundreds of hectares in size. The difference between the cost of a tractor pass in smaller vineyards covering only a few kilometres of rows compared with the possibility of thousands of kilometres of rows, can influence the reality of deciding to do multiple tasks in one pass and the level of consequence in the form of labour and fuel costs for larger vineyards to choose to utilise another tractor pass. However, many of these considerations can also be caused by more granular variables that are not within these models, such as composting, soil depth and clay content. The inability to consider different activitiies at multiple scales without hte use of multiple models becomes a limiting factor.

% ML: As I read this, I wander whether there was some bias because the group were very attuuned to the data and had a strong preference to efficient management techniques. Im not convinced that this is representative of coonawarra per se.

% This is significant... (refering to below)
As a factor, water emerged as the most interconnected variable for both BNs, with edges connecting it to almost everything either directly or indirectly, illustrating its fundamental role in agriculture. Climate conditions also play a significant role, influencing water availability, quality and disease/pest pressures. Within both of the models, the most significant factors impacting sustainability are water quality, pest and disease management, market pressures, agrochemical use, and climate conditions. These factors were identified as having the highest sensitivity to environmental impact, indicating their crucial roles in the sustainability of the region.

@@

The Coonawarra-specific BNs both support the findings from the literature, with a more structured map of how factors interact compared to the Leximancer summary in Figure \ref{fig:lexi}. The BN provides a detailed view of the relationships between factors, highlighting the central role of water and the interconnectedness of pest management, market pressures, and climate conditions. Additionally, the BNs emphasise economic factors, such as market pressures and grape prices, and includes how they interact within a scope of secondary factors like fuel and electricity use.

@@

\section{Discussion}
% weather or climate?
This investigation highlights the key factors of Australian winegrowing that contribute to sustainability through the lens of BNs created with assistance from industry input. These models have established the intricate, and interconnected nature of various influences on sustainability in Australian winegrowing. A keener understanding of how these factors both interact and affect one another from the perspective of research and industry is illustrated within these models, with a major outcome being the highlighting of water use's critical importance as one of the most interconnected variables. The influence of water use reached almost every variable, with its connection to operational decisions being reflected within its propagation to other factors such as tractor use, disease and pest prevention, as well as management strategies for soil. This finding aligns with water being the lifeblood of agriculture~\cite{chawlaWaterProductivityAgriculture2023}, essential for both the health of the vines and the broader ecosystem, especially within Australia's dry climate~\cite{australianbureauofstatisticsWaterUseAustralian2021}. The connection of variables that were linked to operational decisions such as tractor passes, fuel and electricity use illustrated both a way in which the reduction of water use, affects from changes in water quality, climate and disease can influence them. And, how different causal relationships between nodes can be a result of different management perspectives. These differences and relations help to broaden the consideration of factors when making decision in vineyards with regards to how the impact the environment.

@@

The relationships of nodes associated with operational and management  decisions illustrated how different drivers from nodes several edges away from environmetnal impact could still be driving forces through their causal influence of the primary contributing factors. Furthermore, extenral considerations such as market pressures, the demand for high-quality grapes through factors such as grape prices, were found to have an impact on the environment through how they influenced internal responses, such as the use of more water to increase yields as a response to price increases. The economic drivers of some of the decisions would be a great compliment in determinig how the priority of these decision might be made alongside their environmental concerns, and poses a potential avenue for further research.
